---
title: "Pins"
bibliography: "bibliography.bib"
nocite: |
     @*
author: "Robert W. Walker"
format: 
   html:
     self-contained: true
     self-contained-math: true
     code-fold: true
     code-copy: hover
     code-block-border-left: true
     df-print: paged
     embed-resources: true
     fig-format: retina
     footnotes-hover: true
     html-math-method: katex
     mainfont: Playfair Display
     toc: true
     toc-title: Navigation
     theme: simplex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE, warning=FALSE)
```

Creating the pins for this repository.

```
board <- board_temp(versioned = TRUE)
```


## The Data

```{r}
library(tidyverse)
Files <- dir()
Files
```


## A Look at the Data

```{r}
library(DT)
Files.df <- data.frame(Files=Files) %>% mutate(link=paste0(paste0('<a href=',paste0("\"https://github.com/robertwwalker/DADMStuff/raw/master/",Files,"\"", sep=""),' > ', sep=""),Files,'</a>',sep=""))
datatable(Files.df, escape = FALSE)
```
I want to make these files accessible.  Apparently, the way to do this is `pins`.

I am going to put the pins on github.  The repository will be: `https://raw.githubcontent.com/robertwwalker/DADMStuff/pins/`.  Here is a list of the pins.

```{r}
dir("pins/")
```

## How to use them?

First, we have to call `library(pins)`; I will also require the `tidyverse`.

```{r}
#| code-fold: show
library(pins)
library(tidyverse)
board <- board_url("https://raw.githubusercontent.com/robertwwalker/DADMStuff/master/pins/")
board %>% pin_list()
```

Acquiring a single dataset is easiest with `pin_read`.  In this case, let me grab the `EPL` data named in quotations.  Having grabbed it, I will plot it.

```{r}
#| code-fold: show
library(ggrepel)
board %>% pin_read("EPL") %>% 
  ggplot() + 
  aes(y=Points, x=Wage.Bill.milGBP, label=Team) + 
  geom_label_repel() + 
  theme_minimal() +
  labs(x="Wage Bill", y="Points", title="Points and Wages")
```



# References

```{r}
knitr::write_bib(names(sessionInfo()$otherPkgs), file="bibliography.bib")
```

