---
title: "tidyTuesday meets Spending on Children"
author: "A Template"
date: "9/15/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Spending on Kids

First, let me import the data.

```{r}
kids <- read.csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-09-15/kids.csv')
# kids <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-09-15/kids.csv')
```

Now let me summarise it and show a table of the variables.

```{r}
summary(kids)
```

A table of the variables.  [The definitions are best found here.](https://jrosen48.github.io/tidykids/articles/tidykids-codebook.html)

```{r}
table(kids$variable)
```

It is *very* tidy.  It is probably better shown after a pivot.  50 states, the District of Columbia, and 20 years gives us 1,020 observations.  Let me show it wide. 

```{r}
Big.Wide <- pivot_wider(kids, id_cols = c(state,year), names_from = "variable", values_from = c("raw","inf_adj","inf_adj_perchild"))
Big.Wide
```


# My brief plan

I recently came across a `geofacet` for R.  I want to use it to plot a little bit of this data.  If you want to get a head start, try `install.packages("geofacet", dependencies=TRUE)`.  You can google `geofacet` to get an idea of what a `geofacet` plot is.  I will build one on the fly using a couple of tidy tools: filter, mutate, and joins and then put it together.

```{r}
library(geofacet)
```

From the vignette, we have a working example.  

```{r}
ggplot(state_ranks, aes(variable, rank, fill = variable)) +
  geom_col() +
  coord_flip() +
  theme_bw() +
  facet_geo(~ state)
```

```{r}
State.Temp <- state_ranks %>% filter(variable=="education")
Final.Data <- left_join(kids, State.Temp, by = c('state' = 'name')) %>% filter(variable.x == "PK12ed")
ggplot(Final.Data, aes(x=year, y=inf_adj_perchild, color = inf_adj_perchild)) +
  geom_line() +
  theme_void() +
  scale_color_viridis_c() +
  facet_geo(~ state.y) + labs(title="Pre-K through 12 Education Spending per Child", subtitle="Inflation adjusted", color="Spend per Child") -> My.Plot
My.Plot
```

# Getting Real Fancy

`gganimate` makes animation an easy addition to basic ggplots.  The same developer is also responsible for the very accessible `patchwork` for combining multiple ggplots.  I will use the ability to reveal a transition in this animation.

```{r}
library(gganimate)
My.Plot + transition_reveal(year)
```

